# In-Class: September 19, 2018


## Part 1 

1. The data set at rnf6080.dat records hourly rainfall at a certain location in Canada, every day from 1960 to 1980. First, we need to load the data set into R using the command read.table(). Use the help function to learn what arguments this function takes. Once you have the necessary input, load the data set into R and make it a data frame called rain.df.
```{r}
#read.table('rnf6080.dat') -> changed to comment to prevent large output in PDF

rain.df <- read.table('rnf6080.dat') #Read in dataset and set it to variable rain.df
```
2. How many rows and columns does rain.df have? (If there are not 5070 rows and 27 columns, something is wrong; check the previous part to see what might have gone wrong in the previous part.)
```{r}
dim(rain.df) #dim gives us our dimensions (5070x27)
```
3. What are the names of the columns of rain.df?
```{r}
colnames(rain.df) #Returns the current column names of our dataset
```
4. What is the value of row 5, column 7 of rain.df?
```{r}
rain.df[5,7] #Pulls value at (5,7)
```
5. Display the second row of rain.df in its entirety.
```{r}
rain.df[2,] #Pulls each value of row 2
```
6. Explain what the command below does by running it on your data and examining the object. (You may find the display functions head() and tail() useful here.) Is it clear now what the last 24 columns represent?

```{r}
names(rain.df) <- c("year","month","day",seq(0,23)) #Gives text names to the first 3 columns and numbers the remaining ones
names(rain.df) #Confirms that we set out names correctly above
head(rain.df) 
tail(rain.df)
#Displays the begining and the end of our data set
```
**It has changed the names of the columns of the table where the first column is "year", followed by "month", "day"" and then the subsequent columns from are 0-243 which represents the hours of the day (24 hours in a day)**

7. Create a new column in the data frame called daily, which is the sum of the rightmost 24 columns. With this column, create a histogram of the values in this column, which are supposed to be daily rainfall values. What is wrong with this picture?

```{r}
#Create a new column "daily" which is the sum of the rightmost 24 columns
rain.df$daily= apply(rain.df[,4:27], 1, sum)
#Create histogram and label axis and changed color of bars
hist(rain.df$daily, col = "#660066", xlab = "Daily Rain", main = "Daily Rain in Canada")
```

**This histogram shows data points <0 which do not make practal sense given the nature of our data set (rainfall).  This is skewing the graph and making it impossible to analyze properly.**


8. Create a new data frame rain.df.fixed that takes the original and fixes it for the apparent flaw you have discovered. Having done this, produce a new histogram with the corrected data and explain why this is more reasonable.

```{r}
#Clear any values that were less than zero from the dataset
rain.df[rain.df < 0] <- 0
#Create a new column "daily" which is the sum of the rightmost 24 columns
rain.df$daily = apply(rain.df[,4:27], 1, sum)
#Create histogram and label axis and changed color of bars to pink
hist(rain.df$daily, col = "#ff99cc", xlab = "Daily Rain", main = "Daily Rain in Canada")
```

** Now with the values <0 removed we only have either positive rainfall or no rainfall at all.  We are unsure of the large data points however we are not given units of measurement, therefore we cannot definitively pull them out.**